### 2020-01-27 report.tex
initial

Chronological backtracking \cite{DBLP:conf/lpar/2013, DBLP:conf/sat/MohleB19, DBLP:conf/sat/NadelR18} addresses this inefficiency by heuristically backtracking fewer levels up the trail after learning a clause. However, chronological backtracking breaks some traditional invariants of the SAT solver, namely that literals are supposed to appear on the trail in monotonically increasing order by decision level, and can be complex to implement. Additionally, it does not allow the search to have as much flexibility in deciding the order of literals to assign, since leaving a large portion of the trail intact "in a sense" forces decisions originally used to construct that portion to be made. Some empirical evidence also suggests that it is only beneficial when used sparingly (e.g., when the number of levels to be backtracked over is greater than 100).

@@

\subsection{Comparison to Chronological Backtracking}
As mentioned earlier, chronological backtracking also achieves some of the theoretical speedups that trail saving achieves. One advantage of chronological backtracking is that it does not require any caching of the trail or accessing such a cache during propagation; it simply leaves a vast portion of the trail intact rather than backtracking all the way to the conflict level. Chronological backtracking will also speed up propagations slightly more than trail saving, because it leaves more literals assigned at once which leads to clauses being detected as unit more quickly. One disadvantage that chronological backtracking has compared to trail saving is that it doesn't allow the search as much flexibility. Because chronological backtracking leaves a large portion of the trail assigned, it will behave similarly to how the solver would behave if it was to force the decisions that do occur within this large portion of literals, whereas trail saving does not force any decisions. Another disadvantage is that, because propagation is not performed at every decision level, literals that are already assigned do not have a chance to move their decision levels up and this could potentially effect the lbd score or length of the learnt clause that gets generated at the next conflict. An example of this is when the solver was supposed to backtrack over a literal that would end up being unit propagated at level 0, but stays assigned at its current decision level when chronological backtracking is used. Additionally, chronological backtracking breaks the invarant that all literals appear on the trail in monotonically increasing order by decision level, which makes its implementation slightly more complicated.

>>>>

Chronological backtracking \cite{DBLP:conf/lpar/2013, DBLP:conf/sat/MohleB19, DBLP:conf/sat/NadelR18} addresses this inefficiency by heuristically backtracking fewer levels up the trail after learning a clause. However, chronological backtracking breaks some traditional invariants of the SAT solver, namely that literals are supposed to appear on the trail in monotonically increasing order by decision level, and can be complex to implement. Additionally, it does not allow the search to have as much flexibility in deciding the order of literals to assign, since leaving a large portion of the trail intact "in a sense" forces decisions originally used to construct that portion. Some empirical evidence also suggests that it is only beneficial when used sparingly (e.g., when the number of levels to be backtracked over is greater than 100).

@@

\subsection{Comparison to Chronological Backtracking}
As mentioned earlier, chronological backtracking also achieves some of the theoretical speedups that trail saving achieves. One advantage of chronological backtracking is that it does not require any caching of the trail or accessing such a cache during propagation; it simply leaves a vast portion of the trail intact rather than backtracking all the way to the conflict level. Chronological backtracking will also speed up propagations slightly more than trail saving, because it leaves more literals assigned at once which leads to clauses being detected as unit more quickly.

One disadvantage that chronological backtracking has compared to trail saving is that it doesn't allow the search as much flexibility to jump out of local neighbourhoods, as the common variable selection heuristics tend to depend on \cite{DBLP:conf/dac/MoskewiczMZZM01, DBLP:conf/sat/2015, DBLP:conf/sat/LiangGPC16}. Because chronological backtracking leaves a large portion of the trail assigned, it will behave similarly to how the solver would behave if it was to force the decisions that do occur within this large portion of literals, whereas trail saving does not force any decisions. As variable selection heuristics become more refined, it is likely that adhering to them will become more crucial to performance.

Another disadvantage is that, because propagation is not performed at every decision level, literals that remain assigned during chronological backtracking (but otherwise would have been unassigned) do not have a chance to move their decision levels up. An example of this is when the solver was supposed to backtrack over a literal that would end up being unit propagated at level 0, but stays assigned at its current decision level when chronological backtracking is used. This could potentially effect the lbd score or length of the learnt clause that gets generated at the next conflict. Additionally, chronological backtracking breaks the invarant that all literals appear on the trail in monotonically increasing order by decision level, which makes its implementation slightly more complicated.
